### 2024-02-07 products/manuscript/Manuscript.qmd
Project Part 1

---
title: ""
subtitle: ""
author: Mutsa Nyamuranga
date: "02/01/2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

# Summary/Abstract

{{< pagebreak >}}

# Introduction

## General Background Information

Atherosclerosis, a chronic inflammatory disease of the arterial wall, remains a leading cause of cardiovascular morbidity and mortality worldwide. Despite significant advancements in understanding its pathophysiology, effective therapeutic interventions are still in demand. Recent research has shed light on the potential role of dietary β-carotene and its conversion to vitamin A in modulating atherosclerosis progression and resolution.

Several studies have demonstrated the beneficial effects of β-carotene in delaying atherosclerosis progression in experimental models, along with its association with reduced cardiovascular risk in humans. Moreover, genetic variants in the BCO1 gene, responsible for converting β-carotene to vitamin A, have been linked to alterations in circulating carotenoid levels and plasma lipid profiles in individuals. However, the precise mechanisms underlying these effects, particularly regarding the interplay between β-carotene, BCO1 activity, and atherosclerosis progression or resolution, remain unclear.

## Description of data and data source
The data I will be using is based on:

Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.

The data consists of measurements of circulating β-carotene and retinoic acid levels, along with relevant biomarkers and clinical characteristics, collected from a cohort of individuals with varying degrees of atherosclerosis progression and resolution. Genetic information on BCO1 activity, including genetic variants associated with BCO1 activity, may also be included in the dataset. The data was collected through blood sample analysis and genetic testing.

```{r}
# Load necessary packages
library(renv)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths

# Load the dataset (replace "file_path" with the actual file path)
# Load necessary package
library(readxl)

# Read Excel file
rawdata <- read_excel("~/Desktop/MADA-course/Nyamuranga-MADA-project/data/raw-data/elife-87430-data1-v1.xlsx")
```

## Questions/Hypotheses to be addressed

  The question I want to answer using the data is whether there is a correlation between circulating β-carotene levels, BCO1 genetic variants, and the progression or resolution of atherosclerosis. Specifically, I am interested in examining how these factors relate to plasma lipid profiles (e.g., cholesterol, triglycerides) and immune cell populations, particularly Treg levels, in individuals with atherosclerosis. 
  The main outcome of interest will be the degree of atherosclerosis progression or resolution, potentially measured by imaging techniques or clinical assessments of plaque burden. Additionally, plasma lipid profiles and immune cell populations, particularly Treg levels, would be of interest as secondary outcomes. The specific predictors of interest would include circulating β-carotene levels, genetic variants associated with BCO1 activity, plasma lipid profiles, and immune cell populations.
  The data analysis will likely involve statistical methods such as correlation analysis to assess the relationships between predictors and outcomes. Regression modeling could be used to examine the predictive value of circulating β-carotene levels and BCO1 genetic variants on atherosclerosis progression or resolution, controlling for potential confounding variables. Further exploration may involve visualization techniques to identify patterns and trends in the data.

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("products","assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References