### 2024-04-09 R/analysis-code/modeling.qmd
Project 4

---
title: "Model Fitting"
author: "Mutsa Nyamuranga"
date: "2024-03-29"
output: html_document
---

# Introduction

# Set Up

## Load Necessary Packages
First we will load the data and additional packages necessary for the
assigniment.

```{r}
library(readr) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(ggplot2) # for plots
library(gtsummary)# for summary tables
library(patchwork) #for combine plots
library(tidymodels)
library(recipes)
library(parsnip)
```

## Load Data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
figdata <- here::here("data","processed_data","processeddata.rds")
#load data
explorfigdata <- readRDS(figdata)
body_weights <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")
```

# Modeling

Here, I conduct a model fitting to assess the relationship of the
outcome of interest with the other variables. Tidymodels provides the
functions necessary for modeling and preprocessing data. The recipe
function specifies the data preprocessing steps. In this case, the
outcome variable (Y) is defined as the response variable, and all other
variables are considered predictors. step_dummy converts categorical
predictors into dummy variables, step_center centers numeric predictors
around their mean, and step_scale scales numeric predictors to have unit
variance.

## First model fit
```{r}
# fit linear model using height as outcome, weight as predictor

linCat <- lm(BodyWeight ~ Category, body_weights)

# place results from fit into a data frame with the tidy function
linCattable <- broom::tidy(linCat)

#look at fit results
print(linCattable)

# save fit results table  
table_linCat = here("results", "resulttable1.rds")
saveRDS(linCattable, file = table_linCat)
```

## Second model fit
```{r}
# Load required library
library(strucchange)

# Fit a linear regression model with treatment variables
timeseries <- lm(BodyWeight ~ Week + I(Week^2) + I(Week^3) + Category, data = body_weights)

# Summary of the model
summary(timeseries)

# place results from fit into a data frame with the tidy function
timeseriestbl <- broom::tidy(timeseries)

#look at fit results
print(timeseriestbl)

# save fit results table  
seriestable_file2 = here("results", "resulttable2.rds")
saveRDS(timeseriestbl, file = seriestable_file2)
```
## Third Model Fit
```{r}
# Define the recipe
body_recipe <- recipe(BodyWeight ~ ., data = body_weights) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())
```

```{r}
# Split the data into training and testing sets
set.seed(123) # for reproducibility
body_split <- initial_split(body_weights, prop = 0.8, strata = BodyWeight)
body_train <- training(body_split)
body_test <- testing(body_split)

# Define the linear regression model specification
linear_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

# Create a workflow that incorporates the recipe and the model
body_workflow <- workflow() %>%
  add_model(linear_spec) %>%
  add_recipe(body_recipe)
```

The initial_split function divides the dataset into training and testing
sets. The prop argument specifies the proportion of the data to allocate
to the training set (80% in this case), and the strata argument ensures
that the split is stratified based on the outcome variable (Y), which
helps maintain the balance of categories in both sets.

The linear_reg function specifies the linear regression model.
set_engine("lm") selects the linear model engine (ordinary least squares
regression), and set_mode("regression") sets the mode of the model to
regression, indicating that it predicts a continuous outcome (Y).

The workflow function creates a modeling workflow. This workflow
combines the preprocessing steps defined in the recipe with the
specified model. The add_recipe function adds the recipe to the
workflow, and the add_model function adds the linear regression model.

```{r}
body_weights$BodyWeight <- as.numeric(as.character(body_weights$BodyWeight))

class(body_weights$BodyWeight)
```

```{r}
# Fit the workflow to the training data
body_fit <- fit(body_workflow, data = body_train)

# Compute predictions on the testing data
body_predictions <- predict(body_fit, new_data = body_test) %>%
  bind_cols(body_test)
```

The fit function fits the workflow to the training data. This step
applies the preprocessing steps defined in the recipe to the training
data and then fits the linear regression model to the preprocessed data.

The predict function generates predictions from the fitted model using
the testing data. This step applies the preprocessing steps (from the
recipe) to the testing data and then predicts the outcome variable (Y)
using the fitted linear regression model.

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

```{r}
# Compute RMSE and R-squared for the model using all predictors
all_metrics <- body_predictions %>%
  metrics(truth = BodyWeight, estimate = .pred)

all_rmse <- all_metrics %>%
  filter(.metric == "rmse") %>%
  pull(.estimate)

all_r_squared <- all_metrics %>%
  filter(.metric == "rsq") %>%
  pull(.estimate)
```

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

Finally, we print the computed RMSE and R-squared values to evaluate the
model's performance. These metrics provide insights into how well the
linear regression model fits the data and predicts the outcome variable.

```{r}
# Print RMSE and R-squared for the model using all predictors
cat("Model using all predictors:\n")
cat("RMSE:", all_rmse, "\n")
cat("R-squared:", all_r_squared, "\n")

body_predictions
tidy(body_fit)
```

# Results
The predictive model demonstrated favorable performance, with an RMSE of 1.564 and an R-squared value of 0.954. The low RMSE indicates minimal deviation between predicted and actual body weight values, suggesting the model's predictive accuracy. Furthermore, the high R-squared value indicates that the model explains approximately 95.4% of the variance in body weight, indicating strong explanatory power. Interpretation of model coefficients revealed insights into the relative importance of predictors in influencing body weight. interpretation may be necessary depending on the specific context and goals of the analysis.