### 2024-04-25 R/eda-code/exploratory-analysis.qmd
Part 5

---
title: "Data Exploration"
author: "Mutsa Nyamuranga"
date: "2024-02-23"
output: html_document
---

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyr)
library(forcats)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
figdata <- here::here("data","processed_data","processeddata.rds")
#load data
explorfigdata <- readRDS(figdata)
```

```{r}
explorfigdata
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
print(explorfigdata)

summary_table <- explorfigdata %>%
  tbl_summary() %>%
  add_stat(label = "Count", statistic = list(all_continuous() ~ "{n}")) %>%
  add_stat(label = "Mean", statistic = list(all_continuous() ~ "{mean}"))
```

# save to file
summarytable_elife = here("results", "summarytable.rds")
saveRDS(explorfigdata, file = summarytable_elife)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Here, I have created code that develops a bar graph to visualize the distribution of body weight in each condition. This will show the essential volume of body weight throughout the 16 weeks. This will not not be reliable for inferences but essential for drawing a baseline.

## Total Weights

```{r}
# Convert columns to numeric if needed
explorfigdata <- mutate_all(explorfigdata, as.numeric)

# Remove the "Week" variable
explorfigdata1 <- select(explorfigdata, -Week)

# Reshape the data and calculate sum for each condition
explorfigdata_long <- explorfigdata1 %>%
  gather(condition, value) %>%
  filter(!is.na(value)) %>%  # Remove NA values if any
  group_by(condition) %>%
  summarise(sum_value = sum(value))

# Create Bar Graph for Visualization
bar1 <- ggplot(explorfigdata_long, aes(x = fct_reorder(condition, desc(sum_value)), y = sum_value, fill = condition)) +
  geom_col() +
  scale_fill_manual(values = c("red", "blue", "blue", "red", "red", "blue")) +  # Fill colors
  scale_color_manual(values = c("red", "blue", "blue", "red", "red", "blue")) +  # Border colors
  labs(title = "Total Body Weight for Conditions",
       x = "Condition", y = "Sum of Weights")

plot(bar1)
# Saving the plot as an image file
figure_file = here("results","condition_distro.png")
ggsave(filename = figure_file, plot=bar1) 
```

I color coded for male and female stratified conditions. Here, we see that the beta-carotene condition for males has the highest overall total body weight, while the baseline female condition has the lowest weight.

## Weight over time

```{r}
# Reshape data to long format
library(tidyr)
body_weights_long <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")

summary(body_weights_long)
str(body_weights_long)

# Plot the data
library(ggplot2)
ggplot(body_weights_long, aes(x = Week, y = BodyWeight, color = Category)) +
  geom_line() +
  labs(title = "Comparison of Body Weight Changes Across Categories",
       x = "Week", y = "Body Weight") +
  theme_minimal()
figure_file2 = here("results", "figures", "category_distro.png")
ggsave(filename = figure_file2, plot=bar1) 
```

With some final exploration. We look at the change in body weight over the 16 weeks. We see that males start and finish with higher body weights as expected. Of note, we see that the control and Beta-Carotene condition for female mice ends with the lowest body weights.