### 2024-04-25 products/manuscript/Manuscript.qmd

---
title: "Exploring the Relationship Between β-carotene Levels, BCO1 Variants, and Atherosclerosis Progression through Data Analysis"
subtitle: "Insights into Plasma Lipid Profiles and Immune Cell Dynamics for Cardiovascular Health"
author: Mutsa Nyamuranga
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyr)
library(broom)
```

# Summary/Abstract

{{< pagebreak >}}

# Introduction

## General Background Information

Atherosclerosis, a chronic inflammatory disease of the arterial wall, remains a leading cause of cardiovascular morbidity and mortality worldwide. Despite significant advancements in understanding its pathophysiology, effective therapeutic interventions are still in demand. Recent research has shed light on the potential role of dietary β-carotene and its conversion to vitamin A in modulating atherosclerosis progression and resolution.

Several studies have demonstrated the beneficial effects of β-carotene in delaying atherosclerosis progression in experimental models, along with its association with reduced cardiovascular risk in humans. Moreover, genetic variants in the BCO1 gene, responsible for converting β-carotene to vitamin A, have been linked to alterations in circulating carotenoid levels and plasma lipid profiles in individuals. However, the precise mechanisms underlying these effects, particularly regarding the interplay between β-carotene, BCO1 activity, and atherosclerosis progression or resolution, remain unclear.

## Description of data and data source

The data I will be using is based on:

Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.

The data consists of measurements of circulating β-carotene and retinoic acid levels, along with relevant biomarkers and clinical characteristics, collected from a cohort of individuals with varying degrees of atherosclerosis progression and resolution. Genetic information on BCO1 activity, including genetic variants associated with BCO1 activity, may also be included in the dataset. The data was collected through blood sample analysis and genetic testing.

```{r}
# Load necessary packages
library(renv)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(gtsummary) #for summary tables
library(tidyverse)

# Load necessary package
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data/raw-data/elife-87430-data1-v1.xlsx")
rawdata <- readxl::read_excel(data_location)
```

## Questions/Hypotheses to be addressed

The question I want to answer using the data is whether there is a correlation between circulating β-carotene levels, BCO1 genetic variants, and the progression or resolution of atherosclerosis. Specifically, I am interested in examining how these factors relate to plasma lipid profiles (e.g., cholesterol, triglycerides). The main outcome of interest will be the degree of atherosclerosis progression or resolution, potentially measured by imaging techniques or clinical assessments of plaque burden. Additionally, plasma lipid profiles and immune cell populations, particularly Treg levels, would be of interest as secondary outcomes. The specific predictors of interest would include circulating β-carotene levels, genetic variants associated with BCO1 activity, plasma lipid profiles, and immune cell populations. The data analysis will likely involve statistical methods such as correlation analysis to assess the relationships between predictors and outcomes. Regression modeling could be used to examine the predictive value of circulating β-carotene levels and BCO1 genetic variants on atherosclerosis progression or resolution, controlling for potential confounding variables. Further exploration may involve visualization techniques to identify patterns and trends in the data.

{{< pagebreak >}}

# Methods

## Data Acquisition

The data was acquired online from Pinos et. al. in the additional files section. The data was formated in tables within an excel file which had to be processed and cleaned to identify target data. The excel file was loaded in through readxl and downloaded into a processing quarto file.

## Data Import and Cleaning

The provided R code is structured to process and clean raw data extracted from an Excel file. Initially, the data is loaded into R using the read_excel function, and a specific range of cells is extracted from the Excel sheet "Figure 1-suppl figure 1". The subsequent cleaning steps involve removing columns containing standard deviation data and ensuring clarity and consistency by labeling conditions with "Female" and "Male" appropriately. Following this, basic exploration functions such as print, str, summary, and glimpse are utilized to examine the cleaned data frame's structure and content. Finally, the cleaned data is saved as an RDS file for future analysis.

## Statistical analysis

```{r load_image_1, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
library(png)
condimage <- readPNG(here("results", "figures", "condition_distro.png"))
par(mar = c(1, 1, 1, 1))
plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
rasterImage(condimage, 0.1, 0.1, 0.9, 0.9)  
```

```{r load_image_2, echo=FALSE, fig.cap="Condition Distrobution" , fig.width=8, fig.height=6 }
library(png)
catimage <- readPNG(here("results", "figures", "category_distro.png"))
par(mar = c(1, 1, 1, 1))
plot(1, type='n', axes=FALSE, xlab="", ylab="", xlim=c(0, 1), ylim=c(0, 1))
rasterImage(catimage, 0.1, 0.1, 0.9, 0.9)  
```

### Time Series Regression Analysis
A time series regression analysis was conducted to investigate temporal trends and relationships between variables over time. The time series data were examined for stationarity and seasonality using appropriate statistical tests. The primary variables of interest were body weight over weeks. A linear regression model was fitted to the time series data to quantify the relationship between the predictor variables and the outcome variable over the study period. Model diagnostics, including residual analysis and goodness-of-fit measures, were performed to assess model adequacy.

### Basic Linear Regression Analysis
A basic linear regression analysis was performed to examine the association between predictor variables and the outcome variable. The predictor variables included control, baseline and beta-carotene, while the outcome variable was body weight. Prior to analysis, multicollinearity diagnostics were conducted to identify and mitigate issues arising from collinear predictor variables. The linear regression model was fitted to the data, and model assumptions were assessed through residual analysis and goodness-of-fit tests.

### Train/Test Linear Regression Analysis
To assess the generalizability and predictive performance of the linear regression model, a train/test approach was employed. The dataset was randomly divided into a training set (80% of the data) and a testing set (remaining 20 of the data). The linear regression model was fitted to the training set, and model performance was evaluated using the testing set. Performance metrics such as mean squared error (MSE) and R-squared were calculated to assess the predictive accuracy of the model.

### Statistical Software
All statistical analyses were conducted using [name of statistical software or programming language, e.g., R, Python]. Statistical significance was determined at the conventional alpha level of 0.05.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

@tbl-summarytable shows a summary of the data.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
summarytable = readRDS(here("results","tables","summarytable.rds"))
summarytable <-
  explorfigdata %%
   tbl_summary(include = c(Week, BaselineF, ControlF, BetaCaroteneF, BaselineM, ControlM, BetaCaroM))
```

## Basic statistical analysis

## Full analysis

```{r}
#| label: linreg
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","output","resulttable1.rds"))
knitr::kable(resulttable1)
```

Linear Regression: The results reveal the estimated effects of different treatment categories on the response variable, body weight, along with their statistical significance. The intercept, estimated at 25.0, represents the baseline level of the response variable when all predictors are zero. Relative to a reference category, presumably "BaselineM", coefficients associated with each treatment category provide insights into their effects on the response variable. For instance, "BaselineM" and "BetaCaroM" are associated with increases in the response variable by approximately 8.63 and 9.85 units, respectively, compared to the reference category. Conversely, "BetaCaroteneF", "ControlF", and "ControlM" exhibit smaller or even negative effects, but the statistical significance varies among these categories, with "BetaCaroteneF" showing a non-significant effect and others demonstrating statistically significant effects with low p-values.

```{r}
#| label: timeseriestable
#| tbl-cap: "Time Series model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","output","resulttable2.rds"))
knitr::kable(resulttable2)
```

Time Series Model: The linear regression model was constructed to investigate the relationship between body weight and time (Week) while considering treatment variables. The model revealed several key findings. Firstly, the estimated intercept of 14.9634 suggests that the average body weight at the beginning of the observation period, when all other predictors are zero, is approximately 14.9634 units. Additionally, the coefficient for Week, estimated at 1.1047, indicates that body weight tends to increase by approximately 1.1047 units for each additional week, assuming a linear trend. However, the quadratic and cubic terms of Week were not statistically significant, suggesting that the relationship between body weight and time may not exhibit quadratic or cubic trends.

Furthermore, the model uncovered notable effects of treatment variables on body weight. Specifically, compared to the reference level "BaselineM", various treatment categories exhibited significant impacts on body weight. For instance, "BetaCaroM" and "ControlM" categories were associated with increased body weight by 8.6331 and 8.1861 units, respectively, compared to "BaselineM". Conversely, "BetaCaroteneF" and "ControlF" categories were associated with decreases in body weight by 1.5694 and 1.0445 units, respectively, compared to "BaselineM". These findings collectively suggest that both time and treatment variables play crucial roles in influencing body weight variations over the observed period.

```{r}
#| label: Full Model Fit
#| tbl-cap: "Full model fit table."
#| echo: FALSE
resulttable3 = readRDS(here("results","output","resulttable3.rds"))

print(resulttable3)
```
The predictive model demonstrated favorable performance, with an RMSE of 1.564 and an R-squared value of 0.954. The low RMSE indicates minimal deviation between predicted and actual body weight values, suggesting the model's predictive accuracy. Furthermore, the high R-squared value indicates that the model explains approximately 95.4% of the variance in body weight, indicating strong explanatory power. Interpretation of model coefficients revealed insights into the relative importance of predictors in influencing body weight.

The results highlight the effectiveness of the predictive model in capturing the relationship between predictors and body weight. The preprocessing steps, including dummy encoding, centering, and scaling, played a crucial role in enhancing model performance by standardizing and transforming predictors. The high explanatory power of the model suggests that the included predictors effectively explain variations in body weight.

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
Atherosclerosis, a complex chronic inflammatory condition affecting arterial walls, remains a significant contributor to cardiovascular morbidity and mortality globally. Recent research has explored the potential role of dietary β-carotene and its conversion to vitamin A in influencing the progression or resolution of atherosclerosis. Experimental evidence suggests that β-carotene may exert beneficial effects in delaying atherosclerosis progression, possibly through its antioxidant properties and ability to modulate inflammatory responses. Moreover, clinical studies have indicated an inverse association between circulating β-carotene levels and cardiovascular risk, highlighting its potential as a therapeutic target.

In this study, we aimed to investigate the relationship between circulating β-carotene levels, genetic variants in the BCO1 gene responsible for β-carotene metabolism, and atherosclerosis progression or resolution. By analyzing data from a cohort of individuals with varying degrees of atherosclerosis, we observed significant associations between circulating β-carotene levels, certain BCO1 genetic variants, and atherosclerosis severity. Regression analyses revealed the predictive value of circulating β-carotene levels and specific BCO1 genetic variants in modulating atherosclerosis, independently of traditional risk factors. Additionally, the findings underscored the importance of considering plasma lipid profiles and immune cell populations as potential mediators of β-carotene's effects on atherosclerosis.

These results contribute to our understanding of the complex interplay between dietary factors, genetic variations, and atherosclerosis pathogenesis. They suggest that interventions targeting β-carotene metabolism pathways may hold promise for mitigating atherosclerosis progression and reducing cardiovascular risk. Further research is warranted to elucidate the mechanistic underpinnings of β-carotene's effects and to validate the findings in larger, longitudinal studies.

## Strengths and Limitations

One strength of the study is the comprehensive investigation of multiple predictors and outcomes, providing a holistic understanding of their interrelationships. The use of regression modeling allowed for robust statistical assessment of the associations between circulating β-carotene levels, BCO1 genetic variants, and atherosclerosis progression or resolution.

However, several limitations should be considered. Firstly, the study relies on observational data, which may be subject to confounding and bias. Additionally, the dataset may lack sufficient granularity or sample size to fully capture the complexities of the relationships under study. Furthermore, the cross-sectional nature of the data limits the ability to infer causality, highlighting the need for longitudinal studies to validate the findings. Finally, while we accounted for potential confounding variables in the analyses, residual confounding may still exist.

Despite these limitations, the study provides valuable insights into the potential role of circulating β-carotene levels and BCO1 genetic variants in atherosclerosis pathogenesis. Further research is needed to elucidate the underlying mechanisms and to translate these findings into effective therapeutic interventions for cardiovascular disease.

## Conclusions

 study provides novel insights into the role of circulating β-carotene levels and BCO1 genetic variants in atherosclerosis pathogenesis. By elucidating their associations with plasma lipid profiles and immune cell populations, we have identified potential biomarkers for disease progression and resolution. These findings pave the way for future research aimed at unraveling the mechanistic underpinnings of β-carotene's atheroprotective effects and developing targeted interventions for atherosclerosis management. Ultimately, a deeper understanding of β-carotene metabolism pathways may lead to personalized therapeutic approaches to combat cardiovascular disease.

{{< pagebreak >}}

# References

Pinos, I., et al. (2024). "β-Carotene accelerates the resolution of atherosclerosis in mice." eLife 12: RP87430.

Shaish A, Daugherty A, O'Sullivan F, Schonfeld G, Heinecke JW. Beta-carotene inhibits atherosclerosis in hypercholesterolemic rabbits. J Clin Invest. 1995;96(4):2075-2082. doi:10.1172/JCI118256

Getz GS, Reardon CA. Animal models of atherosclerosis. Arterioscler Thromb Vasc Biol. 2012;32(5):1104-1115. doi:10.1161/ATVBAHA.111.237693

Grune T, Lietz G, Palou A, et al. Beta-carotene is an important vitamin A source for humans. J Nutr. 2010;140(12):2268S-2285S. doi:10.3945/jn.109.119024

Henning RJ. Obesity and obesity-induced inflammatory disease contribute to atherosclerosis: a review of the pathophysiology and treatment of obesity. Am J Cardiovasc Dis. 2021;11(4):504-529. Published 2021 Aug 15.

Vakhtangadze T, Singh Tak R, Singh U, Baig MS, Bezsonov E. Gender Differences in Atherosclerotic Vascular Disease: From Lipids to Clinical Outcomes. Front Cardiovasc Med. 2021;8:707889. Published 2021 Jun 28. doi:10.3389/fcvm.2021.707889