### 2024-05-04 R/analysis-code/modeling.qmd
Final Edits
Cleaned up Maunscripts, ReadMe files and deleted unnessary files

---
title: "Model Fitting"
author: "Mutsa Nyamuranga"
date: "2024-03-29"
output: html_document
---

# Introduction

# Set Up

## Load Necessary Packages
First we will load the data and additional packages necessary for the
assigniment.

```{r}
library(readr) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(ggplot2) # for plots
library(gtsummary)# for summary tables
library(patchwork) #for combine plots
library(tidymodels)
library(recipes)
library(parsnip)
library(broom.mixed) # for converting bayesian models to tidy tibbles
library(dotwhisker)  # for visualizing regression results
library(jtools) # Load jtools
```

## Load Data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
figdata <- here::here("data","processed_data","processeddata.rds")
#load data
explorfigdata <- readRDS(figdata)
body_weights <- pivot_longer(explorfigdata, -Week, names_to = "Category", values_to = "BodyWeight")
```

# Modeling

Here, I conduct a model fitting to assess the relationship of the
outcome of interest with the other variables. Tidymodels provides the
functions necessary for modeling and preprocessing data. The recipe
function specifies the data preprocessing steps. In this case, the
outcome variable (Y) is defined as the response variable, and all other
variables are considered predictors. step_dummy converts categorical
predictors into dummy variables, step_center centers numeric predictors
around their mean, and step_scale scales numeric predictors to have unit
variance.

## First model fit

```{r}
# Define the recipe
body1_recipe <- recipe(BodyWeight ~ Category, data = body_weights) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())

# Split data into training and testing sets
set.seed(123)  # for reproducibility
body1_split <- initial_split(body_weights, prop = 0.75, strata = BodyWeight)
body1_train <- training(body1_split)
body1_test <- testing(body1_split)

str(body1_test)
str(body1_train)

# Define the linear regression model specification
linear_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

# Create a workflow that incorporates the recipe and the model
body1_workflow <- workflow() %>%
  add_recipe(body1_recipe) %>%
  add_model(linear_spec)
```
The fit function fits the workflow to the training data. This step
applies the preprocessing steps defined in the recipe to the training
data and then fits the linear regression model to the preprocessed data.

The predict function generates predictions from the fitted model using
the testing data. This step applies the preprocessing steps (from the
recipe) to the testing data and then predicts the outcome variable (Y)
using the fitted linear regression model.

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

```{r}
# Convert BodyWeight to numeric if it's stored as character
body1_train$BodyWeight <- as.numeric(body1_train$BodyWeight)

# Convert BodyWeight to numeric if it's stored as character
body1_test$BodyWeight <- as.numeric(body1_test$BodyWeight)

# Fit the workflow to the training data
linCat <- fit(body1_workflow, data = body1_train)

# Compute predictions on the testing data
body1_preds <- predict(linCat, new_data = body1_test) %>%
  bind_cols(body1_test)
```

```{r}
# Compute RMSE and R-squared for the model
all_metrics <- body1_preds %>%
  metrics(truth = BodyWeight, estimate = .pred)

all_rmse <- all_metrics %>%
  filter(.metric == "rmse") %>%
  pull(.estimate)

all_r_squared <- all_metrics %>%
  filter(.metric == "rsq") %>%
  pull(.estimate)
```

```{r}
# Print RMSE and R-squared for the model
cat("Model:\n")
cat("RMSE:", all_rmse, "\n")
cat("R-squared:", all_r_squared, "\n")

all_rmse
all_r_squared
tidy(linCat)
```
The RMSE is 5.77076

### Model Tuning
So, I will tune the model using cross-validation

The model performance is evaluated using the
cross-validation technique employing 10 folds for the two main models.
This technique involves dividing the training data into 10 subsets and
iteratively using 9 of them for model fitting while reserving the
remaining 10% for evaluation. This process is repeated 10 times,
ensuring that each subset serves as both training and evaluation data.

```{r}
# Train the model using train function from caret package
set.seed(234)  # for reproducibility
folds1 <-vfold_cv(body1_train, v=10)
folds1
```
Subsequently, an object for resampling is constructed using the
'workflow' function from tidymodels. This function combines
pre-processing, modeling, and post-processing instructions into a single
entity, streamlining the analysis pipeline.

```{r}
#Resampling using workflow for the model with only DOSE as predictor
linCat2 <- 
	workflow() %>%
	add_model(linear_spec) %>%
  add_formula(BodyWeight ~ Category)%>%
	fit_resamples(folds1)

collect_metrics(linCat2)
```
Using cross-validation, the RMSE increases to 6.0090840.

```{r}
#| message: false
#| warning: false
# Creating a data-frame with observed and predicted values from the model with `Category` as the predictor
lin1_fit <- linCat %>%
  predict(body1_train) %>%
  bind_cols(body1_train)

#I added a label column and dubbed it 'Simple Linear model'
lin1_fit$label <- rep("Simple Linear Model")

# Create the ggplot figure to graph the predictive values vs the observed value for the three models
p1 <- ggplot(
  lin1_fit, aes(x = BodyWeight, y = .pred, color = label, shape = label)) +
  geom_point(size=2) +
   scale_color_manual(values = c("#9467bd", "#ff9896", "#17becf"))+
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +  # Adding a 45-degree line
  labs(x = "Observed Values", y = "Predicted Values", color = "Model", shape = "Model") +
  xlim(0, 50) + 
  ylim(0, 50)+
  theme_bw()

# Viewing the plot
p1
```
Based on the results of the initial model fitting generating a RMSE of 5.77076 and the cross-validation generating a RMSE increase to 6.0090840, we can conclude that this first model has over-fit the data.

With this, I will move on to other models to find a RMSE that works best for assessing the data. 
```{r}
figure_file3 = here("results", "figures", "SimpleLinearPlot.png")
ggsave(filename = figure_file3, plot= p1) 
```

```{r}
# Summary of the model
summary(linCat)

# place results from fit into a data frame with the tidy function
fullmodel1 <- broom::tidy(linCat)

#look at fit results
print(fullmodel1)

# save fit results table  
model1_file = here("results", "output", "resulttable1.rds")
saveRDS(fullmodel1, file = model1_file)
```

## Second model fit
```{r}
# Load required library
library(strucchange)

# Fit a linear regression model with treatment variables
timeseries <- lm(BodyWeight ~ Week + I(Week^2) + I(Week^3) + Category, data = body_weights)

# Summary of the model
summary(timeseries)

# place results from fit into a data frame with the tidy function
timeseriestbl <- broom::tidy(timeseries)

#look at fit results
print(timeseriestbl)

# save fit results table  
seriestable_file2 = here("results", "resulttable2.rds")
saveRDS(timeseriestbl, file = seriestable_file2)

# Convert the results into a data frame
results <- data.frame(
  term = c("(Intercept)", "Week", "I(Week^2)", "I(Week^3)", 
           "CategoryBaselineM", "CategoryBetaCaroM", 
           "CategoryBetaCaroteneF", "CategoryControlF", "CategoryControlM"),
  estimate = c(14.963391564, 1.104685173, 0.035042256, -0.002315347,
               8.629167500, 8.633072223, -1.569384093, -1.044471462, 8.186054854),
  std.error = c(0.6518010412, 0.2519033316, 0.0292788770, 0.0009802486,
                0.4469748457, 0.4325930512, 0.4325930512, 0.4325930512, 0.4325930512)
)

# Calculate the confidence intervals
results$lower_ci <- results$estimate - 1.96 * results$std.error
results$upper_ci <- results$estimate + 1.96 * results$std.error

# Plot a forest plot
library(ggplot2)

ggplot(results, aes(x = estimate, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci)) +
  labs(title = "Forest Plot of Regression Coefficients",
       x = "Estimate",
       y = "Term") +
  theme_minimal()
```
```{r}
# Convert BodyWeight to numeric
body_weights$BodyWeight <- as.numeric(body_weights$BodyWeight)

# Remove observations with missing values
complete_data <- body_weights[!is.na(body_weights$BodyWeight),]

# Obtain predicted values for complete data
predicted_time_complete <- predict(timeseries, newdata = complete_data)

# Calculate residuals
residuals_complete <- complete_data$BodyWeight - predicted_time_complete

# Calculate RMSE
rmse_complete <- sqrt(mean(residuals_complete^2))

# Print RMSE
print(rmse_complete)
```
The RMSE for this model is 1.210414.

This shows a high level of accuracy in predicting values, however, we can assume that this model may also be over fitting due to an increase level of nuance within the data being taken in to consideration.

## Third Model Fit
```{r}
# Define the recipe
body_recipe <- recipe(BodyWeight ~ ., data = body_weights) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors())
```

```{r}
# Split the data into training and testing sets
set.seed(123) # for reproducibility
body_split <- initial_split(body_weights, prop = 0.8, strata = BodyWeight)
body_train <- training(body_split)
body_test <- testing(body_split)

# Define the linear regression model specification
linear_spec <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

# Create a workflow that incorporates the recipe and the model
body_workflow <- workflow() %>%
  add_model(linear_spec) %>%
  add_recipe(body_recipe)
```

The initial_split function divides the dataset into training and testing
sets. The prop argument specifies the proportion of the data to allocate
to the training set (80% in this case), and the strata argument ensures
that the split is stratified based on the outcome variable (Y), which
helps maintain the balance of categories in both sets.

The linear_reg function specifies the linear regression model.
set_engine("lm") selects the linear model engine (ordinary least squares
regression), and set_mode("regression") sets the mode of the model to
regression, indicating that it predicts a continuous outcome (Y).

The workflow function creates a modeling workflow. This workflow
combines the preprocessing steps defined in the recipe with the
specified model. The add_recipe function adds the recipe to the
workflow, and the add_model function adds the linear regression model.

```{r}
body_train$BodyWeight <- as.numeric(as.character(body_train$BodyWeight))

class(body_train$BodyWeight)
```

```{r}
# Fit the workflow to the training data
body_fit <- fit(body_workflow, data = body_train)

# Compute predictions on the testing data
body_predictions <- predict(body_fit, new_data = body_test) %>%
  bind_cols(body_test)

body_predictions$BodyWeight <- as.numeric(as.character(body_predictions$BodyWeight))
```

The fit function fits the workflow to the training data. This step
applies the preprocessing steps defined in the recipe to the training
data and then fits the linear regression model to the preprocessed data.

The predict function generates predictions from the fitted model using
the testing data. This step applies the preprocessing steps (from the
recipe) to the testing data and then predicts the outcome variable (Y)
using the fitted linear regression model.

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

```{r}
# Compute RMSE and R-squared for the model using all predictors
all_metrics <- body_predictions %>%
  metrics(truth = BodyWeight, estimate = .pred)

all_rmse <- all_metrics %>%
  filter(.metric == "rmse") %>%
  pull(.estimate)

all_r_squared <- all_metrics %>%
  filter(.metric == "rsq") %>%
  pull(.estimate)
```

The metrics function computes evaluation metrics for the model
predictions. Here, we calculate the root mean squared error (RMSE) and
the R-squared value to assess the model's performance. The computed
metrics are then extracted and stored in the variables all_rmse and
all_r_squared, respectively.

Finally, we print the computed RMSE and R-squared values to evaluate the
model's performance. These metrics provide insights into how well the
linear regression model fits the data and predicts the outcome variable.

```{r}
# Print RMSE and R-squared for the model using all predictors
cat("Model using all predictors:\n")
cat("RMSE:", all_rmse, "\n")
cat("R-squared:", all_r_squared, "\n")

body_predictions
tidy(body_fit)
```

```{r}
# Summary of the model
summary(body_fit)

# place results from fit into a data frame with the tidy function
fullmodel <- broom::tidy(body_fit)

#look at fit results
print(fullmodel)

# save fit results table  
model2_file = here("results", "output", "resulttable3.rds")
saveRDS(fullmodel, file = model2_file)
```

# Results
The predictive model demonstrated favorable performance, with an RMSE of 1.564 and an R-squared value of 0.954. The low RMSE indicates minimal deviation between predicted and actual body weight values, suggesting the model's predictive accuracy. Furthermore, the high R-squared value indicates that the model explains approximately 95.4% of the variance in body weight, indicating strong explanatory power. Interpretation of model coefficients revealed insights into the relative importance of predictors in influencing body weight. interpretation may be necessary depending on the specific context and goals of the analysis.